---
title: "Data preprocessing"
output: 
  pdf_document:
    toc: T
author: Marco Galliani    
---
## Settings
```{r}
rm(list = ls())
```


## The dataset
The statistical units are the bombs exploded in the first 24 hours of the bombing of London, on September 7th, 1940. For each bomb the following informations are provided:
- Order
- Time 
- Location
- Type of bomb (IB: Incendiary Bomb, EB: Explosive Bomb, COB: Crude Oil Bomb)
- Damage or other details ("Damage.or.other.details..All.dimensions.at.in.ft.unless.stated.")           )

We got data 843 bombs.

Loading the original dataset, downloaded from [here](http://bombsight.org/data/sources/)
```{r}
library(readxl)

data_path <- "../data/September 7, 1940_ first night of the Blitz.xlsx"

bomb_data <- read_excel(path = data_path,
                        col_types = c("numeric", "date", "text", "text", "text")
                        )

otro_data <- read.csv(file = "../data/September 7, 1940_ first night of the Blitz - SEPT 7, ALL REPORTS.csv")

colnames(bomb_data) <- c("Order", "Time", "Location", "Type.of.bomb", "Damage.or.other")
```

## Preprocessing
- fixing the wrong date imported by the excel
```{r}
library(lubridate)

date(bomb_data$Time) <- "1940-9-7"
```

- getting Latitude and Longitude from the addresses (geocoding)

Done using this online tool: [geoapify](https://www.geoapify.com/tools/geocoding-online)
```{r}
# splitting data since it only accept datasets of size up to 500 rows
first_batch <- bomb_data[1:500, 3]
second_batch <- bomb_data[501:nrow(bomb_data), 3]

write.csv(first_batch, "temp_data/first_batch.csv")
write.csv(second_batch, "temp_data/second_batch.csv")
```

```{r}
geocoded_first_batch <- read.csv("temp_data/geocoded_first_batch.csv")
geocoded_second_batch <- read.csv("temp_data/geocoded_second_batch.csv")

geocoded_data <- rbind.data.frame(geocoded_first_batch, geocoded_second_batch)
rm(first_batch, second_batch, geocoded_first_batch, geocoded_second_batch)

bomb_data[, c("lat", "lon")] <- geocoded_data[, c("lat", "lon")]
rm(geocoded_data)
```

- getting the district where the bomb fell
```{r}
library(terra)
library(tidyterra)
library(ggplot2)

# loading data for the map of London
london_spat_vect <- vect("../data/London-data/London_Borough_Excluding_MHW.shp")

# projecting the longitude and latitude values on the reference system used by the geocoding service
newcrs <- "+proj=longlat +datum=WGS84"
london_spat_vect <- terra::project(london_spat_vect, newcrs)

# getting the districts related to each (lon,lat) value
bomb_data$district <- extract(london_spat_vect, bomb_data[,c("lon","lat")])$NAME
```

- cleaning duplicates in type of bomb
```{r}
bomb_data$Type.of.bomb <- as.factor(bomb_data$Type.of.bomb)

levels(bomb_data$Type.of.bomb) <- list(IB = c("IB", "Ib", "IBIB"),
                                       EB = c("EB", "EB ", "eb", "Eb", "High Explosive Bomb"),
                                       EB.and.IB = c("EB & IB", "EB &IB", "EB &IB", "IB & EB", "IB and EB"),
                                       Magnesium.Flare = c("Magnesium Flare"),
                                       Shrapnel = c("Shrapnel"),
                                       COB = c("COB", "Crude oil bomb"),
                                       Crashed.Aircraft = c("Crashed aircraft"),
                                       Unknow.enemy.action = c("Unknown enemy action"))

```
- cleaning useless variables generated
```{r}
bomb_data$Order <- NULL
knitr::kable(head(bomb_data))

write.csv(bomb_data, file = "../data/geocoded_bomb_data.csv")
```
